---
title: "Homework 4_1"
author: "Jyoti Chaudhary"
date: "February 24, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sas7bdat)
```

#Solution 1.a


```{r }
# Fitting the MLR model using LM function

xr16054 <- read.csv(paste(getwd(),"/xr16054.csv",sep = ""),header=T)

xr16054_fit <- lm(formula = xr16054$Salary ~ xr16054$GPA + xr16054$Activities, data = xr16054)

summary(xr16054_fit)


```

The MLR equation from above R output is:

Y-hat = 24.3092 + 3.8416 * X1 +  1.6810 * X2

where 
Beta0 = 24.3092
Beta1 = 3.8416
Beta2 = 1.6810

X1 = GPA
X2 = Activities
Y-hat = Predicted Salary

Interpretation of regression Coefficients:
a. For a unit increase in GPA and a constant Activities, the starting salary will increase by 3.8416 units.
b. For a unit increase in number of Activities with constant age, the starting salary will increase by 1.6810 units. 
c. As per the regression equation, if GPA is zero and number of activities is zero, the starting salary equals 24.3092. This interpretation is invalid as the parameter values in this case are outside the range of test data.

# Solution (1.b)

For X1 = 3.6 and X2 = 3, the estimated salary will be:

Y-predicted = 24.3092 + 3.8416 * 3.6 +  1.6810 * 3 = 43.18196

#Hence, Dave's estimated starting salary = 43.18196


# Solution (1.c)

```{r }

# Below is the code to calculate SSE = e'e where e = (y - y-hat) = Observed y minus predicted y

# code to prepare a 10 by 1 matrix for regressor GPA and ACTIVITIES
matrix_X1 <- matrix(xr16054$GPA)
matrix_X2 <- matrix(xr16054$Activities)
# Code to prepare a 10 by 1 matrix
matrix_Beta0 <- matrix(rep(1,length(xr16054$Activities))) 
# code to combine columns of all the above matrices
X <- cbind(matrix_Beta0, matrix_X1, matrix_X2) 
# Matrix for the response variable
Y <- matrix(xr16054$Salary)

Beta_hat <- solve(t(X)%*%X)%*%(t(X)%*%Y) 

# As e = y - (X * Beta-hat)
e <- Y -(X%*%Beta_hat)
# SSE = e'e
sse <- t(e) %*% (e)

# Standard Error = sqrt(SSE / n - (k + 1))
SE <- sqrt(sse/(10-(2+1)))
SE

```

From above output, it can be seen that Standard Error for the model = 1.448. This value can also be found from R output: "Residual standard error: 1.448" 


#Solution 3.a


```{r }

# Fitting the MLR model

StreetVN <- read.sas7bdat(paste(getwd(),"/STREETVN.sas7bdat",sep = ""))

StreetVN_fit <- lm(formula = StreetVN$earnings ~ StreetVN$age + StreetVN$hours, data = StreetVN)

summary(StreetVN_fit)


```

# Solution 3.c
From above model fit summary output, the regression equations =

Y-hat = -20.352 + 13.350 * X1 + 243.714 * X2
where 
Beta0 = -20.352
Beta1 = 13.350
Beta2 = 243.714

X1 = Age
X2 = Hours
Y-hat = Predicted Earnings

Interpretation of regression Coefficients:
a. For a unit increase in age, keeping all other regressors constant, the Earnings of Street Vendors increase by a factor of 13.350

b. For a unit increase in hours worked per day, keeping all other regressors constant, the Earnings of Street Vendors increase by a factor of 243.714

c. for age = zero and hours worked per day equals zero, the estimated earnings will be -20.352. This interpretation has no practical significance since the age and hours worked per day are outside the range of test data used for model fitting and is not even practically feasible as age and hours worked cannot be zero.

