---
title: "Exam3_Jyoti_Chaudhary"
author: "Jyoti Chaudhary"
date: "May 7, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(MASS)
library(MKmisc)
library(ResourceSelection)
library(car)
library(gdata)

```

# Problem 1

a)

```{r }
data1 <- read.csv(paste(getwd(),"/Test_S18_Data.csv",sep = ""),header=T)

mdl1= glm ( data1$Owner ~ data1$Income, data = data1, family = binomial)

summary(mdl1) ##### PARAMETER ESTIMATES 

# The coefficients obtained are:

mdl1$coefficients


```

The fitted model is:

$$\hat {y} = \hat {\pi} = \frac {1}{1+e^{-(-6.2289 + 0.1319x)}} = \frac {1}{1+e^{6.2289 - 0.1319x}}$$ 
where x = Income

If the model is good at 5% level?

```{r }

anova(mdl1, test="Chisq")

hoslem.test(data1$Owner, fitted(mdl1), g=8) 

vcov(mdl1)

# Test of model significance
0.1319/sqrt(0.003687)
(0.1319/sqrt(0.003687))^2
qchisq(.95,df = 1)



```

Null deviance= 26.921 Residual deviance= 17.634, it shows a significant improvement over null model. As D/df = 0.98 which is close to 1.00, this indicates that the model is adequate and good fit for the data. Also, the p-value (=0.002308) is very low which further indicates that model is adequate.

We can also do HL-test to test goodness of fit which shows Chi-square = 3.4905 with 6 degrees of freedom and p-value
=  0.7452. Hence it is good fit to the data.

Also, as (0.1319/sqrt(0.003687))^2 > qchisq(.95,df = 1), the model is adequate and significant.

# b)
Find and interpret the odds ratio for "Income".

```{r }
OR=exp(coef(mdl1)[2])
OR
```

This odds ratio implies that for every unit increase in speed, the odds of hitting the target decrease by 1.75%.

Using odds ratio we can say that for every 1 unit increase of income, the odds of owning the device increase by 100x(1.141101 - 1) = 14.1%.


# c)
Predict the probability of ownership if income is $40,000. Also nd a 98% condence interval for the true probability.

```{r }

nwdt=with(data1, data.frame(Income=40.0)) ##### NEW DATA POINT
nwdt

pct=0.98

nwdt2=subset(cbind(nwdt,predict(mdl1,newdata=nwdt, type="link", se=TRUE)),select = -c(residual.scale))
nwdt2

nwdt3=within(nwdt2,{PredictedProb <- plogis(fit)
LL <- plogis(fit - (qnorm((1+pct)/2) * se.fit))
UL <- plogis(fit + (qnorm((1+pct)/2) * se.fit))})
nwdt3


```


#d) 

```{r }

data1$Income2 <- data1$Income^2
mdl2= glm ( data1$Owner ~ data1$Income + data1$Income2 , data = data1, family = binomial)


summary(mdl2) ##### PARAMETER ESTIMATES 

anova(mdl2, test="Chisq")

```

The deviance has NOT significantly reduced due to quadratic term with a high p-value of 0.474182. Hence its evident that the interaction term is not required in the model. 

To test Beta3 = 0 vs Beta3 not = 0,
Difference in Deviance = 17.634 - 17.121 =0.513 which is lesser than chi-square(0.1,1) = 2.706. Hence the quadratic term is not significant at 10% level.


#Problem 3

a)

```{r }

data2 <- read.csv(paste(getwd(),"/Test_S18_Data_2.csv",sep = ""),header=T)

# fitting linear regression data

mod3 = lm(Yt ~ t , data = data2)

# plotting residuals against time

plot(data2$t, mod3$res, ylab="Residuals", xlab="time", main="Residual Plot over Time - Model-1", type="o") 
abline(0, 0)   

# the plot indicates positive autocorrelation


# Dubin watson test to determine correlation

library(car)

durbinWatsonTest(mod3, max.lag=1, simulate=TRUE, reps=10000, method="normal", alternative="positive")

```

n = 20, k = 1, dw = 0.308722, alpha = 0.05, dL = 1.201, dU = 1.411
As dw < dL , reject H0, which means there is evidence to support the conclusion that the residuals are positively autocorrelated.



# b)

cochrane orcutt method


```{r }

##### CALCULATING RHOHAT USING FORMULA ############

res3=resid(mod3)
n= length(res3)
rho = sum(res3[1:(n-1)]*res3[2:n])/sum(res3^2)
rho


library(DataCombine)

data3=slide(data=data2, Var="Yt", TimeVar="t",  NewVar="Yt_1", slideBy = -1,
           keepInvalid = FALSE, reminder = TRUE)

pst2=slide(data=data3, Var="t", TimeVar="t",  NewVar="t_1", slideBy = -1,
           keepInvalid = FALSE, reminder = TRUE)

pst2

pst2$yprime= pst2$Yt - rho*pst2$Yt_1
pst2$xprime= pst2$t - rho*pst2$t_1

pst2

mod5 = lm(yprime ~ xprime , data = pst2)
mod5
summary(mod5)

dw2=durbinWatsonTest(lm(yprime ~ xprime,data = pst2) , max.lag=1, alternative="positive")
dw2

```


n = 19, k = 1, dw = 1.778269, alpha = 0.05, dL = 1.180, dU = 1.401

As dw > dU , which means that there is no auto-correlation present among errors. So we conclude that there is no problem with autocorrelated errors in the transformed model. The CochraneOrcutt method has been effective in removing the autocorrelation.


