---
title: "Homework8"
author: "Jyoti Chaudhary"
date: "April 16, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

#Problem 1.a
Fit a logistic regression model to the response variable y.

```{r }
data1 <- read.csv(paste(getwd(),"/Data_HW_8_1.csv",sep = ""),header=T)

mdl1= glm ( y ~ x, data = data1, family = binomial)

summary(mdl1) ##### PARAMETER ESTIMATES 

# The coefficients obtained are:

mdl1$coefficients


```

The fitted model is:

$$\hat {y} = \hat {\pi} = \frac {1}{1+e^{-(6.07 - 0.0177x)}} = \frac {1}{1+e^{-6.07 + 0.0177x}}$$ 

#(1.b)
Does the model deviance indicate that the logistic regression model from part a is adequate?

```{r }

anova(mdl1, test="Chisq")

```

The Deviance = 20.364 with df = 23. As D/df = 0.8 which is close to 1.00, this indicates that the model is adequate and good fit for the data. Also, the p-value (=0.0001597) is very low which further indicates that model is adequate.

#(1.c) 
Provide an interpretation of the parameter beta1 in this model.

```{r }
OR=exp(coef(mdl1)[2])
OR
```

This odds ratio implies that for every unit increase in speed, the odds of hitting the target decrease by 1.75%.

#(1.d)
Expand the linear predictor to include a quadratic term in target speed. Is there any evidence
that this quadratic term is required in the model?
```{r }

mdl2= glm ( y ~ x + x^2, data = data1, family = binomial)

anova(mdl1, test="Chisq")

```

There is no difference in the deviance of the two models and hence there is no need for the quadratic term.


#(2.a)
Fit a logistic regression model to the data.

```{r }

data2 <- read.csv(paste(getwd(),"/Data_HW_8_2.csv",sep = ""),header=T)

mdl3= glm ( y ~ x1 + x2, data = data2, family = binomial)

summary(mdl3) ##### PARAMETER ESTIMATES 

# The coefficients obtained are:

mdl3$coefficients


```

The fitted model is:

$$\hat {y} = \hat {\pi} = \frac {1}{1+e^{-(-7.047 + 7.382e-05x1 + 0.9879x2)}} = \frac {1}{1+e^{7.047 - 7.382e-05x1 - 0.9879x2}}$$ 


```

#(2.b)
Does the model deviance indicate that the logistic regression model from part a is adequate?

```{r }

anova(mdl3, test="Chisq")

```

The Deviance = 21.08 with df = 17. As D/df = 1.24 which is greater than 1.00, this indicates that the model fits the data, however model fit can be improved by further analysis.Also, the p-value (=0.01506) is low which further indicates that model is adequate.


#(2.c) 
Interpret the model coefficients beta1 and beta2.

```{r }

OR=exp(coef(mdl3))
OR
```

The above odds ratio implies that: 

- For a unit increase in the income, the odds of purchasing a new vehicle increase by 0.0074%. 
- For a unit increase in the age of family's oldest vehicle, the odds of purchasing a new vehicle increase by 268.5% which implies the odds are more than approx 2.5 times.

#(2.d)
What is the estimated probability that a family with an income of $45,000 and a car that is 5
years old will purchase a new vehicle in the next 6 months?

```{r }

pred2=subset(data.frame(data2,predict(mdl3, se.fit=TRUE, type='response')),select = -c(residual.scale))

nwdt=with(pred2, data.frame(x1=45000,x2=5))  # NEW DATA POINT
nwdt

pct=0.95

nwdt2=subset(cbind(nwdt,predict(mdl3,newdata=nwdt, type="link", se=TRUE)),select = -c(residual.scale))
nwdt2

nwdt3=within(nwdt2,{PredictedProb <- plogis(fit)
LL <- plogis(fit - (qnorm((1+pct)/2) * se.fit))
UL <- plogis(fit + (qnorm((1+pct)/2) * se.fit))})
nwdt3

```

We can see from the above result that the predicted probability is 0.77 of purchasing a new vehicle.

#(2.e)
Expand the linear predictor to include an interaction term. Is there any evidence that this term
is required in the model?


```{r }

mdl4= glm ( y ~ x1*x2, data = data2, family = binomial)

summary(mdl4) ##### PARAMETER ESTIMATES 

anova(mdl4, test="Chisq")

```

The deviance has significantly reduced due to interaction term with a p-value of 0.033. As D/df = 1.03, this indicates that the model is adequate and good fit for the data. Hence its evident that the interaction term is required in the model. 


#(2.f)
If income goes up by $1000 in model of part (a) while age remain fixed, how much the odds of buying change.

```{r }

OR=exp(coef(mdl3)[2])
OR

```

From the odds ratio we can determine that if the income goes up by $1000, the odds of purchasing a new vehicle increase by 7.4%.

#(2.g)

Find approximate 95% confidence intervals on the model parameters for the logistic regression
model from part a.

```{r }

confint(mdl3, level=0.95) 

```


