---
title: "Homework7"
author: "Jyoti Chaudhary"
date: "April 8, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reshape2) 

```

#Problem 1
This is the case of two way Anova without interaction. Here Factor A = portfolio and Factor B = Advisor.
k = 3, b = 4.

```{r }
data2 <- read.csv(paste(getwd(),"/XR12095.csv",sep = ""),header=T)

  
library(reshape2) 
data3 <- melt(data2)
data3

mdl1 <- aov(value ~ Advisor + variable, data = data3) 

# Now lets compare the portfolios
TukeyHSD(mdl1, conf.level = 0.95, "variable")
plot(TukeyHSD(mdl1,"variable"))

```

From the output p-values and the plot, we can determine that there is no signficant difference between portfolios at 5% level. 


```{r }

with(data3, pairwise.t.test(x=value, g=variable, p.adjust="none")) ## FISHER'S LSD 

```


None of the portfolio comparisons are statistically significant. Hence we conclude that there is no significant difference between the expected annual rate of return between portfolios at 5% level. 


# Problem 2:

This is one-way anova test. 

```{r }
data1 <- read.csv(paste(getwd(),"/XR12096.csv",sep = ""),header=T)

Stacked_data1 <- stack(data1)

mdl2 <- aov(values ~ ind, data = Stacked_data1) 

```

Lets compare the brands using 0.025 level.

```{r }

TukeyHSD(mdl2, conf.level = 0.975, "ind")

```

Looking at the p-value, we can further determine that at 0.025 level, there is no statistically significant difference between the breaking strength of different brand types.

```{r }

with(Stacked_data1, pairwise.t.test(x=values, g=ind, p.adjust="none")) ## FISHER'S LSD 

```

Except for brandD-brandA, none of the other portfolio comparisons are statistically significant. For brandD-brandA, there is statistically significant difference between the breaking strength. For all other brand comparisons, we conclude that there is no significant difference between the breaking strength at 2.5% level. 

# Problem 3:

```{r }
data4 <- read.csv(paste(getwd(),"/XR12099.csv",sep = ""),header=T)

data5 <- melt(data4)

mdl3 <- aov(value ~ Style * variable, data = data5)

summary(mdl3)

TukeyHSD(mdl3,"Style")

```

Using 0.05 level of significance, we can conclude that Type style 1 and type style 2 has no significant difference in the effect on publication readability. 

```{r }

TukeyHSD(mdl3,"variable")

```
Using 0.05 level of significance, we can conclude that Only medium and light darkness levels have statistically significant difference in the effect on publication readability. 

```{r }

TukeyHSD(mdl3,"Style:variable")

```
Using 0.05 level of significance, we can conclude that the interaction effect of style type and darkness level do not have statistically significant difference in the effect on publication readability.

# Problem 4

```{r }

data6 <- read.csv(paste(getwd(),"/LtnSqr.csv",sep = ""),header=T)

mdl4 <- aov(Yield ~ Peanuts + Column + Row, data = data6)


TukeyHSD(mdl4,"Peanuts")

```
Using 0.05 level of significance, we can conclude that the comparison in difference of yield of various peanuts varities is not statistically significant.